---
title: "Comparing_supply_curves"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

August 14th, 2019
Patricia Levi
Compare observed supply curves in historical ERCOT data as well as in my modeled data.
```{r}
library(tidyverse)
library(lubridate)
library(readxl)
library(viridis)
library(gridExtra)
knitr::opts_chunk$set(echo = FALSE) 

base_fol = "/Users/patricia/Documents/Google Drive/stanford/Value of DR Project/"
input_fol = "Data/julia_input/"
default_in_fol = paste0(base_fol,input_fol,"ercot_default/")
```
# What does historical supply curve look like?
```{r include=FALSE}
demand = read_csv(file = paste0(default_in_fol,"ercot_demand_2016.csv"))
demand$Date = mdy_hm(demand$time_CST, tz="US/Central")

supplycurve = merge(demand, medprice, by="Date")
```


```{r include=FALSE}
# load and prep input supply curve data for plotting
# Taken from ercot_gen_inspection.R

dataout = "/Users/patricia/Documents/Google Drive/stanford/Value of DR Project/Data/"
# gendat = read.csv(file = paste0(dataout,"unformatted data/all_cols_complete_generator_listing_ERCOT_012019.csv"))
gendat = read.csv(file = paste0(dataout,"julia_input/ercot_default/complete_generator_correctedGas_3-12-19_alldata.csv"))
plotsfol = "/Users/patricia/Documents/Google Drive/stanford/Value of DR Project/plots/ERCOT_data/"
# plot cumulative supply ####
# REMOVE DR
DRsel = gendat$Fuel == "DR"
gendat1 = gendat
gendat = gendat1[!DRsel,]

ordered = order(gendat[,"VCost"]) # gives the indices of VCost in ascending value
sorted = gendat[,"Capacity"][ordered]
cumsorted = cumsum(sorted)
plotdf <- data.frame(x=cumsorted, 
                     cost=gendat[,"VCost"][ordered],
                     totcost = (gendat[,"VCost"][ordered] + gendat[,"StartCost"][ordered]/264)/1e3, #spread startup costs over 264 hours for full_test
                     Technology=gendat[,"Fuel"][ordered],
                     cap = gendat[,"Capacity"][ordered],
                     GENID = gendat[,"plantUnique"][ordered])
plotdf$loc <- cumsum(plotdf$cap) - plotdf$cap/2
plotdf$cost[which(plotdf$cost==0)] = -0.001

```

```{r}
# find total production by time - approximate with scenario for now
# totdemand = prod_o1 %>%
#   group_by(t) %>%
#   summarise(dem = sum(MWout))
# medprice$t = 24*yday(medprice$Date) + hour(medprice$Date) - 23
# supplycurve = merge(totdemand, medprice, by="t")
ggplot(supplycurve,aes(x=demand, y=medprice)) + geom_point(size = 0.01) +
  # coord_cartesian(ylim = c(0,100)) +
  labs(x = "Demand", y = "Median Price", title = "Historical demand and median price") +
  theme_bw()
```
Zoom ...
```{r}
ggplot(supplycurve,aes(x=demand, y=medprice)) + geom_point(size = 0.01) +
  coord_cartesian(ylim = c(0,100)) +
  labs(x = "Demand", y = "Median Price", title = "Zoom - Historical demand and median price")+
  theme_bw()
```

# What does modeled supply curve look like?
Examining baseline run from May 2nd
```{r include=FALSE}
# get total demand in o1 - could probably be done faster with vdem files...
prod_o1 = read_csv(file = paste0(base_fol,"Data/julia_output/forIAEE_1Pmin/base_noDRfullyear_2019-05-02/","prod_o1.csv"))

totdemand = prod_o1 %>%
  group_by(t) %>%
  summarise(dem = sum(MWout))
rm(prod_o1)
# NB: totdemand is missing first 4 hours and also last few hours of year.

prod_margprice_o1 = read_csv(file = paste0(base_fol,"Data/julia_output/forIAEE_1Pmin/base_noDRfullyear_2019-05-02/","prod_margprice_o1.csv"))
# from fullyear_baseline_analysis.Rmd

prod_margprice_o1 = merge(prod_margprice_o1, totdemand, by = "t", all.x = T)

```

```{r}
ggplot(prod_margprice_o1,aes(x=dem, y=margprice)) + geom_point(size = 0.01) +
  # coord_cartesian(ylim = c(0,100)) +
  labs(x = "Modeled Demand", y = "Marginal generator cost", title = "Observed supply curve") +
  theme_bw()
```

# side by side
<!-- https://sebastiansauer.github.io/two-plots-rmd/ -->
```{r eval=FALSE, include=FALSE}
p1 = ggplot(prod_margprice_o1,aes(x=dem, y=margprice)) + geom_point(size = 0.01) +
  # coord_cartesian(ylim = c(0,100)) +
  labs(x = "Modeled Demand", y = "Marginal generator cost", title = "Modeled supply curve") +
  coord_cartesian(ylim = c(0,100), xlim = c(20000,80000)) +
  theme_bw()
p2 = ggplot(supplycurve,aes(x=demand, y=medprice)) + geom_point(size = 0.01) +
  coord_cartesian(ylim = c(0,100), xlim = c(20000,80000)) +
  labs(x = "Demand", y = "Median Price", title = "Zoom - Historical supply curve")+
  theme_bw()
grid.arrange(p1, p2, ncol = 1)
```
```{r}
p1 = ggplot(prod_margprice_o1,aes(x=dem, y=margprice)) + geom_point(size = 0.01) +
  # coord_cartesian(ylim = c(0,100)) +
  labs(x = "Modeled Demand", y = "Marginal generator cost ($/MWh)", title = "Modeled supply curve") +
  geom_line(aes(x=loc,y=cost),plotdf[!DRsel,], color = "red") +
  coord_cartesian(ylim = c(0,100)) +
  theme_bw()
p2 = ggplot(supplycurve,aes(x=demand, y=medprice)) + geom_point(size = 0.01) +
  geom_line(aes(x=loc,y=cost),plotdf[!DRsel,], color = "red") +
  coord_cartesian(ylim = c(0,100)) +
  labs(x = "Demand", y = "Median Price ($/MWh)", title = "Zoom - Historical supply curve")+
  theme_bw()
grid.arrange(p1, p2, ncol = 2)
```



```{r}
  # testing just a line on top of generator supply curve
# ggplot(plotdf[!DRsel,], aes(x=loc,y=cost,width=cap)) + 
#   #  geom_point(aes(color=Technology)) +
#   #  geom_bar(binwidth = 0,aes(fill=Technology), stat="identity", position="identity") +
#   geom_bar(stat="identity") +
#   geom_line() +
#   labs(x="Cumulative Capacity (MW)",y="Variable Cost ($/MWh)") + 
#   ggtitle(paste("Texas' Electriciy Supply Curve: 2016")) + 
#   theme(legend.position = "bottom") #+ 

# combined plot
ggplot() + 
  geom_line(aes(x=loc,y=cost),plotdf[!DRsel,], color = "red") +
  geom_point(aes(x=dem, y = margprice),prod_margprice_o1, size = 0.01) +
  # coord_cartesian(ylim = c(0,100)) +
  labs(x = "Modeled Demand", y = "Marginal generator cost", title = "Input supply curve + simulated supply curve") +
  coord_cartesian(ylim = c(0,100)) +
  theme_bw()
  
```

```{r}
# compare with simulation
ggplot() + 
  geom_line(aes(x=loc,y=cost),plotdf[!DRsel,], color = "red") +
  geom_point(aes(x=demand, y = medprice),supplycurve, size = 0.01) +
  # coord_cartesian(ylim = c(0,100)) +
  labs(x = "Demand", y = "Marginal generator cost", title = "Input supply curve + historical supply curve") +
  coord_cartesian(ylim = c(0,115)) +
  theme_bw()

```

# examine dual variable of supply=demand
Compare shadow price of constraint (in o1) to demand level (o1) and to the marginal generator cost (o1)....
```{r message=FALSE}
demanddual = read_csv(file = paste0(outputfol,"demanddual.csv"))
# dim(demanddual)
nscenarios = 25
demdf = merge(totdemand,demanddual, by = "t")
```

```{r}
ggplot() + 
  geom_point(aes(x=dem, y = x1*nscenarios),demdf, size = 0.01) +
  geom_line(aes(x=loc,y=cost),plotdf[!DRsel,], color = "red") +
  coord_cartesian(ylim = c(0,115)) +
  labs(x = "Demand", y = "Marginal generator cost", title = "Input supply curve + dual values of supply-demand") +
  # coord_cartesian(ylim = c(0,100)) +
  theme_bw()
```

```{r}
p3 = ggplot(demdf,aes(x=dem, y = x1*nscenarios)) + 
  coord_cartesian(ylim = c(0,100)) +
  geom_point(size = 0.01) + 
  geom_line(aes(x=loc,y=cost),plotdf[!DRsel,], color = "red") +
  labs(x="Demand MW", y = "Shadow price of supply>=demand", title = "Shadow price of supply >= demand") +
  theme_bw()

grid.arrange(p2,p3,ncol = 2)
```
The dual variable of supply = demand looks much more like we might expect.

How does the dual variable compare in terms of distribution?
```{r}
h1 = ggplot(demdf, aes(x=x1*nscenarios)) + geom_histogram(binwidth = 5) +
  coord_cartesian(xlim = c(-10,200)) +
  labs(title = "Simulated price", x = "") +
  theme_bw()
h2 = ggplot(supplycurve,aes(x=medprice)) + geom_histogram(binwidth = 5) +
  coord_cartesian(xlim = c(-10,200)) +
  labs(title = "Historical price", x = "$/MWh")+
  theme_bw()
grid.arrange(h1,h2,ncol = 1)
```

